---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Assuming the user is in the correct working directory, data is loaded like this:
```{r}
data<-read.csv("activity.csv")
```
The data is not processed or transformed.

## What is mean total number of steps taken per day?

The following code produces a histogram of the total number of steps taken each day: 
```{r}
library(plyr)
Total<-ddply(data,~date,summarise,total=sum(steps))
barplot(Total$total, xlab="Date", ylab="Total", main="Total steps per day")
```
The mean and median number of steps taken per day are calculated by this code:
```{r}
ddply(data,~date,summarise,mean=mean(steps, na.rm = TRUE),median=median(steps, na.rm = TRUE))
```

## What is the average daily activity pattern?
This code calculates a time series plot for the average number of steps taken each day (averaged across all days) for each five-minute interval.
```{r}
mean_interval<-ddply(data,~interval,summarise, mean=mean(steps, na.rm = TRUE))
plot(mean_interval, type="l", xlab= "Interval", ylab= "Average number of steps (across all days)")
```
This code determines the interval with the largest average number of steps, averaged across all days:
```{r}
mean_interval$interval[mean_interval$mean == max(mean_interval$mean)]
```

## Imputing missing values
To calcalculate the number of rows featuring 'NA' in the 'steps' column:
```{r}
length(which(is.na(data$steps)))
```

NAs are then replaced with the median value for that particular interval, and a new dataset is produced:

```{r}
library("data.table")
data2<-data.table(data)
data2[,steps := ifelse(is.na(steps), median(steps, na.rm=TRUE), steps), by=interval]
write.csv(data2, "activity_imputed.csv")
```

Calulcating the new mean and median from the new dataset:

```{r}
ddply(data2,~date,summarise,mean=mean(steps, na.rm = TRUE),median=median(steps, na.rm = TRUE))
```

Compare these values to those shown above to demonstrate the effect of imputing missing values.

## Are there differences in activity patterns between weekdays and weekends?

First, a new column is added to the dataset with the day of the week listed:
```{r}
data2$day <- weekdays(as.Date(data2$date))
```
Then these values are replaced with either "Weekend" or "Weekday" as appropriate:
```{r}
data2$day[data2$day == "Sunday"] <- "Weekend"
data2$day[data2$day == "Saturday"] <- "Weekend"
data2$day[data2$day !== "Weekend"] <- "Weekday"
```

